---
title: "Mnemonic Similarity Task (MST)"
output:
  html_document: default
  pdf_document: default
date: "2026-03-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggplot2)
library(readr)
library(stringr)
```

## Load Dataset
```{r}
data_path <- "MST_Data"

task_files <- list.files(data_path,
                         pattern = "_task_",
                         recursive = TRUE,
                         full.names = TRUE)

test_files <- list.files(data_path,
                         pattern = "_test_",
                         recursive = TRUE,
                         full.names = TRUE)

task_conditions <- data.frame(file = task_files) %>%
  mutate(
    condition = case_when(
      str_detect(file, "item_only") ~ "item_only",
      str_detect(file, "task_only") ~ "task_only",
      str_detect(file, "Both_item_task") ~ "Both_item_task"
    )
  )

test_conditions <- data.frame(file = test_files) %>%
  mutate(
    condition = case_when(
      str_detect(file, "item_only") ~ "item_only",
      str_detect(file, "task_only") ~ "task_only",
      str_detect(file, "Both_item_task") ~ "Both_item_task"
    )
  )

table(task_conditions$condition)
table(test_conditions$condition)
```

## Read and Combine All Participant Data
```{r}
read_task <- function(file){

  df <- read_csv(file,
                 show_col_types = FALSE,
                 col_types = cols(.default = col_character()),
                 name_repair = "minimal")

  df$participant <- str_extract(basename(file), "^[0-9]+")

  # Extract experimental condition
  df$condition <- case_when(
    str_detect(file, "item_only") ~ "item_only",
    str_detect(file, "task_only") ~ "task_only",
    str_detect(file, "Both_item_task") ~ "Both_item_task",
    TRUE ~ NA_character_
  )

  if("key_resp_9.rt" %in% names(df))
    df$key_resp_9.rt <- as.numeric(df$key_resp_9.rt)

  if("key_resp_8.rt" %in% names(df))
    df$key_resp_8.rt <- as.numeric(df$key_resp_8.rt)

  return(df)
}


read_test <- function(file){

  df <- read_csv(file, show_col_types = FALSE)

  df$participant <- str_extract(basename(file), "^[0-9]+")

  df$condition <- case_when(
    str_detect(file, "item_only") ~ "item_only",
    str_detect(file, "task_only") ~ "task_only",
    str_detect(file, "Both_item_task") ~ "Both_item_task",
    TRUE ~ NA_character_
  )

  return(df)
}

task_data <- bind_rows(lapply(task_files, read_task))
test_data <- bind_rows(lapply(test_files, read_test))
```

## Clean Encoding Data
```{r}
encoding <- task_data %>%
  select(participant,
         condition,
         image_path,
         key_resp_9.rt,
         key_resp_8.rt,
         key_resp_9.keys,
         key_resp_8.keys) %>%
  mutate(
    rt = case_when(
      
     !is.na(key_resp_9.rt) ~ key_resp_9.rt,
     !is.na(key_resp_8.rt) ~ 3 + key_resp_8.rt,
     TRUE ~ NA_real_
)
  )
```

## Identify Boundary Position (Post / Mid / Pre)
```{r}

encoding <- encoding %>%
  group_by(participant) %>%
  mutate(
    trial = row_number(),
    pos_in_event = ((trial - 1) %% 7) + 1,
    boundary_type = case_when(
      pos_in_event == 1 ~ "Post",
      pos_in_event == 7 ~ "Pre",
      TRUE ~ "Mid"
    )
  ) %>%
  ungroup()
```

## Reaction Time Analysis
```{r}
rt_summary <- encoding %>%
  group_by(condition, boundary_type) %>%
  summarise(
    mean_rt = mean(rt, na.rm = TRUE),
    sd_rt = sd(rt, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

rt_summary
```

## Plot Reaction Times
```{r}
ggplot(encoding, aes(x = boundary_type, y = rt)) +
  geom_boxplot(aes(fill = boundary_type)) +
  facet_wrap(~condition) +
  labs(
    title = "Reaction Time across Event Positions by Condition",
    x = "Event Position",
    y = "Reaction Time (seconds)"
  ) +
  theme_minimal()
```

## Reaction Time mean plot per condition
```{r}
rt_plot <- encoding %>%
  group_by(condition, boundary_type) %>%
  summarise(mean_rt = mean(rt, na.rm=TRUE), .groups="drop")

ggplot(rt_plot, aes(x=boundary_type, y=mean_rt, fill=boundary_type)) +
  geom_bar(stat="identity") +
  facet_wrap(~condition) +
  labs(
    title="Mean Reaction Time by Condition",
    x="Event Position",
    y="Mean Reaction Time"
  ) +
  theme_minimal()
```

## Identify Stimulus Type (Test Phase)
```{r}
test_clean <- test_data %>%
  select(participant, condition, image_path, key_resp_3.keys, key_resp_3.rt) %>%
  mutate(
    stimulus_type = case_when(
      str_detect(image_path, regex("foil", ignore_case = TRUE)) ~ "foil",
      str_detect(image_path, "b\\.jpg$") ~ "lure",
      str_detect(image_path, "a\\.jpg$") ~ "target",
      TRUE ~ NA_character_
    )
  ) %>%
  filter(!is.na(stimulus_type))
```

## Response Distribution
```{r}
response_summary <- test_clean %>%
  group_by(condition, stimulus_type, key_resp_3.keys) %>%
  summarise(n = n(), .groups = "drop")

response_summary
```

## Visualization – Response Counts
```{r}
ggplot(response_summary,
       aes(x = stimulus_type, y = n, fill = key_resp_3.keys)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~condition) +
  labs(
    title = "Participant Responses by Stimulus Type and Condition",
    x = "Stimulus Type",
    y = "Count"
  ) +
  theme_minimal()
```

## Response Accuracy
```{r}
test_clean <- test_clean %>%
  mutate(
    correct = case_when(
      stimulus_type == "target" & key_resp_3.keys == "o" ~ 1,
      stimulus_type == "foil" & key_resp_3.keys == "n" ~ 1,
      stimulus_type == "lure" & key_resp_3.keys == "s" ~ 1,
      TRUE ~ 0
    )
  )
```

## Accuracy Statistics
```{r}
accuracy_summary <- test_clean %>%
  group_by(condition, stimulus_type) %>%
  summarise(
    accuracy = mean(correct, na.rm = TRUE),
    n = n(),
    .groups = "drop"
  )

accuracy_summary
```

## Visualization – Accuracy
```{r}
ggplot(accuracy_summary,
       aes(x = stimulus_type, y = accuracy, fill = stimulus_type)) +
  geom_bar(stat = "identity") +
  facet_wrap(~condition) +
  labs(
    title = "Recognition Accuracy by Condition",
    x = "Stimulus Type",
    y = "Accuracy"
  ) +
  theme_minimal()
```

## Recognition Memory Index (REC)
```{r}
REC_summary <- test_clean %>%
  group_by(condition) %>%
  summarise(
    p_old_target = mean(key_resp_3.keys[stimulus_type=="target"] == "o"),
    p_old_foil = mean(key_resp_3.keys[stimulus_type=="foil"] == "o"),
    REC = p_old_target - p_old_foil
  )

REC_summary
```

## Lure Discrimination Index (LDI)
```{r}
LDI_summary <- test_clean %>%
  group_by(condition) %>%
  summarise(
    p_sim_lure = mean(key_resp_3.keys[stimulus_type=="lure"] == "s"),
    p_sim_foil = mean(key_resp_3.keys[stimulus_type=="foil"] == "s"),
    LDI = p_sim_lure - p_sim_foil
  )

LDI_summary
```

## REC/LDI visualization
```{r}
rec_ldi <- merge(REC_summary, LDI_summary, by="condition")

rec_ldi_long <- rec_ldi %>%
  tidyr::pivot_longer(cols=c(REC,LDI),
                      names_to="metric",
                      values_to="value")

ggplot(rec_ldi_long,
       aes(x=condition,y=value,fill=metric)) +
  geom_bar(stat="identity",position="dodge") +
  labs(
    title="Recognition Memory Metrics by Condition",
    x="Condition",
    y="Score"
  ) +
  theme_minimal()
```
  
## Inferential Test (t-test)
```{r}
encoding %>%
  group_by(condition) %>%
  summarise(
    p_value = t.test(
      rt[boundary_type == "Post"],
      rt[boundary_type == "Mid"],
      na.rm = TRUE
    )$p.value
  )
```