---
title: "Event Boundaries and Recognition Memory: Report 1"
subtitle: "BRSM 2026 — Behavioural Research and Statistical Methods"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    highlight: tango
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
    toc_depth: 3
    number_sections: true
    fig_width: 8
    fig_height: 4
---

```{css, echo=FALSE}
body { font-size: 14px; line-height: 1.7; }
h1   { color: #2E5FA3; border-bottom: 2px solid #2E5FA3; padding-bottom: 5px; }
h2   { color: #3A6FA3; }
h3   { color: #555; }
.note {
  background: #EEF3FB;
  border-left: 4px solid #2E5FA3;
  padding: 10px 14px;
  margin: 10px 0;
  border-radius: 3px;
}
```

---

# Setup

## Install Packages *(run once, then leave eval = FALSE)*

```{r install, eval=FALSE}
install.packages(c("tidyverse", "knitr", "kableExtra"))
```

## Load Libraries

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)   # data wrangling + ggplot2
library(knitr)       # tables in output
library(kableExtra)  # styled tables
```

## Configuration

::: {.note}
**Set `DATA_DIR` to the path of your `MST_Data` folder.**  
The `.Rmd` file and `MST_Data/` folder must be in the same directory (e.g., `MST_Exp/`).  
If they are, leave `DATA_DIR <- "MST_Data"` as-is.
:::

```{r config}
DATA_DIR   <- "MST_Data"      # change if needed

PATHS <- list(
  both      = file.path(DATA_DIR, "Both_item_task",  "both_data"),
  item_only = file.path(DATA_DIR, "item_only",       "item_only_data"),
  task_only = file.path(DATA_DIR, "task_only",       "task_only_data")
)

OUTPUT_DIR <- "outputs"
dir.create(OUTPUT_DIR, showWarnings = FALSE)
```

---

# Introduction

## Background

We experience life as a continuous stream of events, yet our minds naturally divide this stream
into meaningful chunks separated by **event boundaries** — moments when one situation ends
and another begins (Zacks et al., 2007). These transitions fundamentally shape how we encode
and later retrieve episodic memories.

One well-established way to measure the fidelity of long-term memory is the
**Mnemonic Similarity Task (MST)** (Stark et al., 2019). Unlike simple old/new recognition
tasks, the MST presents three types of test stimuli:

- **Targets** — exact repetitions of studied images (tests basic recognition)
- **Lures** — images that are *similar but not identical* to studied items (tests fine-grained
  mnemonic discrimination, thought to rely on hippocampal pattern separation)
- **Foils** — completely novel images never seen before (serves as a false-alarm baseline)

Lures are further organised into **5 similarity bins**, where Bin 1 = most similar to the target
(hardest to distinguish) and Bin 5 = least similar (easiest to distinguish). This graded design
allows researchers to examine how memory fidelity degrades as perceptual interference
increases.

## The Present Study

This study replicates and extends Morse et al. (2023), who found that event boundaries exert
**directional** effects on recognition memory:

- Items encoded **immediately after** a boundary (Post-Boundary) showed *worse* target
  recognition.
- Items encoded **immediately before** a boundary (Pre-Boundary) showed *better* lure
  discrimination.

The core hypothesis is that as an event boundary approaches, the hippocampus is driven into
a high-fidelity encoding state — forming a kind of "snapshot" of the ending event — while
simultaneously creating a brief attentional "blind spot" for the incoming event.

Our version of the experiment differs from Morse et al. in three ways:

1. The encoding question for objects was changed to *"Will it fit in a shoebox?"*
   (replacing *"indoor or outdoor?"*)
2. Events consist of **7 images** per block (40 events total; 280 encoding trials)
3. Test phase: **60 Targets** (20 per position), **60 Lures** (20 per position),
   **30 Foils** — using a three-response format: *old / new / similar*

Three participant groups were tested, varying the nature of the event boundary:

| Group | Boundary Type | n |
|---|---|---|
| **Both** | Item class shift + task shift | 51 |
| **Item-Only** | Item class shift only | 56 |
| **Task-Only** | Task shift only | 53 |

## Key Measures

$$\text{REC} = P(\text{"old"} \mid \text{Target}) - P(\text{"old"} \mid \text{Foil})$$

$$\text{LDI} = P(\text{"similar"} \mid \text{Lure}) - P(\text{"similar"} \mid \text{Foil})$$

Both indices correct for response bias by subtracting the Foil false-alarm rate. A value above
zero means performance is better than chance; below zero means it is worse than the
false-alarm baseline.

---

# Methods

## Data Loading and Organisation (Week 4)

### Parsing the Test File

Each participant contributes one test CSV (150 rows). The relevant columns are:

| Column | Meaning |
|---|---|
| `image_path` | filename encodes stimulus type: `a.jpg` = Target, `b.jpg` = Lure, `Foil/` prefix = Foil |
| `trials.key_resp_3.keys` | Response: `o` = old, `n` = new, `s` = similar |
| `trials.key_resp_3.rt` | Reaction time in seconds |
| `position_of_stimuli` | Boundary position from PsychoPy: `post`, `pre`, `mid`, `none` (Foils) |

```{r parse-test}
parse_test_file <- function(filepath, group_label) {
  raw <- read_csv(filepath, show_col_types = FALSE)
  # Use filename as unique participant ID to avoid numeric ID collisions across groups
  pid <- tools::file_path_sans_ext(basename(filepath))

  raw %>%
    filter(!is.na(image_path)) %>%
    transmute(
      participant  = pid,
      group        = group_label,
      image_path   = str_trim(image_path),
      # Classify stimulus type from filename
      stim_type = case_when(
        str_detect(image_path, regex("Foil", ignore_case = TRUE)) ~ "Foil",
        str_ends(image_path, "b.jpg")                              ~ "Lure",
        str_ends(image_path, "a.jpg")                              ~ "Target",
        TRUE ~ NA_character_
      ),
      # Image class (Object or Scene)
      img_type = case_when(
        str_detect(image_path, regex("Object", ignore_case = TRUE)) ~ "Object",
        str_detect(image_path, regex("Scene",  ignore_case = TRUE)) ~ "Scene",
        TRUE ~ "Unknown"
      ),
      # Decode response keys
      response = case_when(
        `trials.key_resp_3.keys` == "o" ~ "old",
        `trials.key_resp_3.keys` == "n" ~ "new",
        `trials.key_resp_3.keys` == "s" ~ "similar",
        TRUE ~ NA_character_
      ),
      RT_ms = `trials.key_resp_3.rt` * 1000,
      # Map position labels
      boundary_pos = case_when(
        position_of_stimuli == "post" ~ "Post-Boundary",
        position_of_stimuli == "pre"  ~ "Pre-Boundary",
        position_of_stimuli == "mid"  ~ "Mid-Boundary",
        TRUE ~ NA_character_     # Foils have no boundary position
      ),
      boundary_pos = factor(boundary_pos,
                            levels = c("Post-Boundary", "Mid-Boundary", "Pre-Boundary"))
    ) %>%
    filter(!is.na(stim_type))
}
```

### Parsing the Encoding (Task) File

Each participant also has one task CSV (280 encoding trials). The key RT correction
(from the experiment brief): if the participant responded **before** the image disappeared,
`key_resp_9.rt` records the RT; if they responded **after** the image (which was shown for
3 seconds), the total RT = 3 + `key_resp_8.rt`.

```{r parse-task}
parse_task_file <- function(filepath, group_label) {
  raw <- read_csv(filepath, show_col_types = FALSE)
  # Use filename as unique participant ID to avoid numeric ID collisions across groups
  pid <- tools::file_path_sans_ext(basename(filepath))

  # FIX: Some participants always responded before image offset, so PsychoPy
  # never created key_resp_8.rt in their CSV. Add it as NA to prevent crash.
  if (!"trials.key_resp_8.rt" %in% names(raw)) raw$trials.key_resp_8.rt <- NA_real_
  if (!"trials.key_resp_9.rt" %in% names(raw)) raw$trials.key_resp_9.rt <- NA_real_

  raw %>%
    filter(
      !is.na(trials.thisN),
      !str_detect(image_path, regex("practice", ignore_case = TRUE))
    ) %>%
    transmute(
      participant  = pid,
      group        = group_label,
      trial_n      = as.integer(trials.thisN),
      image_path   = str_trim(image_path),
      img_type = case_when(
        str_detect(image_path, regex("Object", ignore_case = TRUE)) ~ "Object",
        str_detect(image_path, regex("Scene",  ignore_case = TRUE)) ~ "Scene",
        TRUE ~ "Unknown"
      ),
      # Correct RT: responded before image = key_resp_9; after = 3 + key_resp_8
      RT_ms = case_when(
        !is.na(`trials.key_resp_9.rt`) ~ as.numeric(`trials.key_resp_9.rt`) * 1000,
        !is.na(`trials.key_resp_8.rt`) ~ (3 + as.numeric(`trials.key_resp_8.rt`)) * 1000,
        TRUE ~ NA_real_
      )
    ) %>%
    arrange(trial_n) %>%
    mutate(
      # 7 items per event: positions 0-6 within each block of 7
      pos_in_event = trial_n %% 7L,
      boundary_pos = case_when(
        pos_in_event == 0 ~ "Post-Boundary",
        pos_in_event == 6 ~ "Pre-Boundary",
        TRUE              ~ "Mid-Boundary"
      ),
      boundary_pos = factor(boundary_pos,
                            levels = c("Post-Boundary", "Mid-Boundary", "Pre-Boundary"))
    )
}
```

### Load All Participants

```{r load-data, message=FALSE, warning=FALSE}
load_group <- function(folder, group_label, type = c("task", "test")) {
  type    <- match.arg(type)
  pattern <- if (type == "task") "MST_task.*\\.csv$" else "MST_test.*\\.csv$"
  parser  <- if (type == "task") parse_task_file else parse_test_file
  files   <- list.files(folder, pattern = pattern, full.names = TRUE)
  map_dfr(files, ~ parser(.x, group_label))
}

# Load test (recognition) data
test_data <- bind_rows(
  load_group(PATHS$both,      "Both",      "test"),
  load_group(PATHS$item_only, "Item-Only", "test"),
  load_group(PATHS$task_only, "Task-Only", "test")
) %>%
  mutate(group = factor(group, levels = c("Item-Only", "Task-Only", "Both")))

# Load encoding (task) data
encoding_data <- bind_rows(
  load_group(PATHS$both,      "Both",      "task"),
  load_group(PATHS$item_only, "Item-Only", "task"),
  load_group(PATHS$task_only, "Task-Only", "task")
) %>%
  mutate(group = factor(group, levels = c("Item-Only", "Task-Only", "Both")))

# Confirm participant counts
test_data %>%
  distinct(participant, group) %>%
  count(group, name = "N_participants") %>%
  kable(caption = "Participants loaded per group") %>%
  kable_styling(bootstrap_options = c("striped", "hover"), full_width = FALSE) %>%
  row_spec(0, bold = TRUE, background = "#2E5FA3", color = "white")
```

## Computing REC and LDI (Week 4)

```{r compute-scores}
# Step 1: Foil false-alarm rates (baseline)
foil_rates <- test_data %>%
  filter(stim_type == "Foil") %>%
  group_by(participant, group) %>%
  summarise(
    p_old_foil = mean(response == "old",     na.rm = TRUE),
    p_sim_foil = mean(response == "similar", na.rm = TRUE),
    .groups = "drop"
  )

# Step 2: Target hit rates and Lure "similar" rates by boundary position
target_rates <- test_data %>%
  filter(stim_type == "Target", !is.na(boundary_pos)) %>%
  group_by(participant, group, boundary_pos) %>%
  summarise(p_old_target = mean(response == "old", na.rm = TRUE),
            .groups = "drop")

lure_rates <- test_data %>%
  filter(stim_type == "Lure", !is.na(boundary_pos)) %>%
  group_by(participant, group, boundary_pos) %>%
  summarise(p_sim_lure = mean(response == "similar", na.rm = TRUE),
            .groups = "drop")

# Step 3: Merge and compute bias-corrected scores
scores <- target_rates %>%
  full_join(lure_rates,  by = c("participant", "group", "boundary_pos")) %>%
  left_join(foil_rates,  by = c("participant", "group")) %>%
  mutate(
    REC = p_old_target - p_old_foil,   # Recognition index
    LDI = p_sim_lure   - p_sim_foil    # Lure discrimination index
  )

# Step 4: Non-Boundary baseline subtraction
non_boundary <- scores %>%
  filter(boundary_pos == "Mid-Boundary") %>%
  select(participant, group, REC_non = REC, LDI_non = LDI)

diff_scores <- scores %>%
  filter(boundary_pos != "Mid-Boundary") %>%
  left_join(non_boundary, by = c("participant", "group")) %>%
  mutate(
    REC_diff = REC - REC_non,
    LDI_diff = LDI - LDI_non
  )
```

---

# Results

## Descriptive Statistics (Week 3)

### Raw Response Rates

```{r raw-response-rates}
raw_rates <- test_data %>%
  filter(!is.na(boundary_pos) | stim_type == "Foil") %>%
  mutate(pos_label = ifelse(stim_type == "Foil", "Foil", as.character(boundary_pos))) %>%
  group_by(group, stim_type, pos_label) %>%
  summarise(
    N        = n(),
    pct_old  = round(mean(response == "old",     na.rm = TRUE), 3),
    pct_new  = round(mean(response == "new",     na.rm = TRUE), 3),
    pct_sim  = round(mean(response == "similar", na.rm = TRUE), 3),
    .groups  = "drop"
  ) %>%
  filter(!(stim_type == "Foil" & pos_label != "Foil"))

raw_rates %>%
  arrange(group, stim_type, pos_label) %>%
  kable(caption = "Table 1. Mean response proportions by stimulus type, boundary position, and group.",
        col.names = c("Group", "Stim Type", "Position", "N",
                      "P(old)", "P(new)", "P(similar)")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2E5FA3", color = "white")
```

### Descriptive Statistics for REC and LDI

```{r desc-stats}
desc <- scores %>%
  group_by(group, boundary_pos) %>%
  summarise(
    N      = n(),
    REC_M  = round(mean(REC, na.rm = TRUE), 3),
    REC_SD = round(sd(REC,   na.rm = TRUE), 3),
    LDI_M  = round(mean(LDI, na.rm = TRUE), 3),
    LDI_SD = round(sd(LDI,   na.rm = TRUE), 3),
    .groups = "drop"
  )

desc %>%
  kable(caption = "Table 2. Descriptive statistics for REC and LDI by group and boundary position.",
        col.names = c("Group", "Boundary Position", "N",
                      "REC Mean", "REC SD", "LDI Mean", "LDI SD")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"), full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2E5FA3", color = "white")

write_csv(desc, file.path(OUTPUT_DIR, "table2_descriptive_REC_LDI.csv"))
```

## Visualisations (Week 3)

### Colour Palette and Theme

```{r theme, include=FALSE}
bp_col <- c("Post-Boundary" = "#E07B54",
            "Non-Boundary"  = "#7EA8BE",
            "Pre-Boundary"  = "#5A9E6F")

theme_mst <- theme_classic(base_size = 12) +
  theme(strip.background = element_blank(),
        strip.text        = element_text(face = "bold", size = 12),
        plot.title        = element_text(face = "bold", size = 13),
        legend.position   = "none")
```

### Figure 1: Distribution of Response Types by Group

Before examining the key memory scores, it is useful to see how participants distributed their
*old*, *new*, and *similar* responses across the three groups.

```{r fig1, fig.cap="Figure 1. Mean proportion of each response type by stimulus type and group. Error bars = ±1 SE across participants.", fig.height=5}
resp_summary <- test_data %>%
  filter(!is.na(response)) %>%
  group_by(participant, group, stim_type, response) %>%
  summarise(prop = mean(!is.na(response)), .groups = "drop") %>%
  group_by(participant, group, stim_type) %>%
  mutate(prop = sum(response == response) / n()) %>%   # re-summarise cleanly below
  ungroup()

resp_clean <- test_data %>%
  filter(!is.na(response)) %>%
  group_by(participant, group, stim_type) %>%
  summarise(
    old     = mean(response == "old"),
    new     = mean(response == "new"),
    similar = mean(response == "similar"),
    .groups = "drop"
  ) %>%
  pivot_longer(cols = c(old, new, similar),
               names_to = "response", values_to = "prop")

resp_se <- resp_clean %>%
  group_by(group, stim_type, response) %>%
  summarise(M = mean(prop), SE = sd(prop)/sqrt(n()), .groups = "drop") %>%
  mutate(response = factor(response, levels = c("old","similar","new")))

fig1 <- resp_se %>%
  ggplot(aes(x = response, y = M, fill = stim_type)) +
  geom_col(position = "dodge", alpha = 0.85, width = 0.65) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE),
                position = position_dodge(0.65), width = 0.2) +
  scale_fill_manual(values = c("Target" = "#2E5FA3",
                                "Lure"   = "#E07B54",
                                "Foil"   = "#7EA8BE"),
                    name = "Stimulus") +
  facet_wrap(~group) +
  labs(x = "Response", y = "Mean Proportion",
       title = "Figure 1. Response Proportions by Stimulus Type and Group") +
  theme_classic(base_size = 12) +
  theme(strip.background = element_blank(),
        strip.text        = element_text(face = "bold"),
        plot.title        = element_text(face = "bold"))

fig1
ggsave(file.path(OUTPUT_DIR, "fig1_response_proportions.png"),
       fig1, width = 9, height = 4, dpi = 300)
```

### Figure 2: Target Hit Rates by Boundary Position

```{r fig2, fig.cap="Figure 2. Target hit rate (proportion 'old' responses) by boundary position and group. Points = individual participants; bar = group mean ± 95% CI."}
fig2 <- scores %>%
  ggplot(aes(x = boundary_pos, y = p_old_target, fill = boundary_pos)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.55, width = 0.45) +
  geom_jitter(width = 0.12, alpha = 0.45, size = 1.3, colour = "grey30") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4,
               colour = "black") +
  scale_fill_manual(values = bp_col) +
  facet_wrap(~group) +
  labs(x = "Boundary Position", y = "Target Hit Rate P(\"old\" | Target)",
       title = "Figure 2. Target Hit Rates by Boundary Position and Group") +
  theme_mst

fig2
ggsave(file.path(OUTPUT_DIR, "fig2_target_hit_rates.png"),
       fig2, width = 9, height = 3.8, dpi = 300)
```

### Figure 3: Lure "Similar" Rates by Boundary Position

```{r fig3, fig.cap="Figure 3. Proportion of 'similar' responses to Lures by boundary position and group. Higher values reflect better mnemonic discrimination."}
fig3 <- scores %>%
  ggplot(aes(x = boundary_pos, y = p_sim_lure, fill = boundary_pos)) +
  geom_boxplot(outlier.shape = NA, alpha = 0.55, width = 0.45) +
  geom_jitter(width = 0.12, alpha = 0.45, size = 1.3, colour = "grey30") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 4,
               colour = "black") +
  scale_fill_manual(values = bp_col) +
  facet_wrap(~group) +
  labs(x = "Boundary Position", y = "Lure CR Rate P(\"similar\" | Lure)",
       title = "Figure 3. Lure Discrimination Rates by Boundary Position and Group") +
  theme_mst

fig3
ggsave(file.path(OUTPUT_DIR, "fig3_lure_sim_rates.png"),
       fig3, width = 9, height = 3.8, dpi = 300)
```

### Figure 4: REC and LDI Scores Side-by-Side

```{r fig4, fig.cap="Figure 4. Bias-corrected REC (left) and LDI (right) scores by boundary position and group. Dashed line at zero = chance/baseline. Diamond = group mean.", fig.height=4}
scores_long <- scores %>%
  select(participant, group, boundary_pos, REC, LDI) %>%
  pivot_longer(cols = c(REC, LDI), names_to = "Measure", values_to = "Score")

fig4 <- scores_long %>%
  ggplot(aes(x = boundary_pos, y = Score, fill = boundary_pos)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_boxplot(outlier.shape = NA, alpha = 0.55, width = 0.45) +
  geom_jitter(width = 0.12, alpha = 0.35, size = 1.0, colour = "grey30") +
  stat_summary(fun = mean, geom = "point", shape = 18, size = 3.5,
               colour = "black") +
  scale_fill_manual(values = bp_col) +
  facet_grid(Measure ~ group) +
  labs(x = "Boundary Position", y = "Score",
       title = "Figure 4. REC and LDI by Boundary Position and Group") +
  theme_mst +
  theme(axis.text.x = element_text(angle = 25, hjust = 1, size = 9))

fig4
ggsave(file.path(OUTPUT_DIR, "fig4_REC_LDI_scores.png"),
       fig4, width = 9, height = 5, dpi = 300)
```

### Figure 5: Encoding Reaction Times — Manipulation Check

A Post-Boundary slowdown in encoding RT is evidence that participants *perceived* the
boundary as a genuine event transition (Morse et al., 2023; DuBrow & Davachi, 2013).

```{r fig5, fig.cap="Figure 5. Mean encoding RT (ms) by boundary position and group. Error bars = ±1 SE. A Post-Boundary slowdown confirms participants experienced the event transitions."}
rt_summary <- encoding_data %>%
  group_by(participant, group, boundary_pos) %>%
  summarise(mean_RT = mean(RT_ms, na.rm = TRUE), .groups = "drop") %>%
  group_by(group, boundary_pos) %>%
  summarise(M  = mean(mean_RT, na.rm = TRUE),
            SE = sd(mean_RT,   na.rm = TRUE) / sqrt(n()),
            .groups = "drop")

fig5 <- rt_summary %>%
  ggplot(aes(x = boundary_pos, y = M, fill = boundary_pos)) +
  geom_col(alpha = 0.85, width = 0.55) +
  geom_errorbar(aes(ymin = M - SE, ymax = M + SE), width = 0.2) +
  scale_fill_manual(values = bp_col) +
  facet_wrap(~group) +
  labs(x = "Boundary Position", y = "Mean Encoding RT (ms)",
       title = "Figure 5. Encoding Reaction Times by Boundary Position and Group") +
  theme_mst

fig5
ggsave(file.path(OUTPUT_DIR, "fig5_encoding_RT.png"),
       fig5, width = 9, height = 3.8, dpi = 300)
```

## Null Hypothesis Testing (Week 6)

We use **one-sample t-tests** to ask whether the mean REC or LDI difference score
(Boundary − Non-Boundary baseline) is significantly different from zero.

- **H₀**: Mean difference score = 0 (boundary position has no effect on memory)
- **H₁**: Mean difference score ≠ 0 (boundary position affects memory)

We apply a **Bonferroni correction** across the four primary tests reported here
(α = 0.05 / 4 = **0.0125**).

> Note: Independent-samples comparisons between boundary conditions (Pre vs. Post) will be
> reported in Report 2 once we have covered comparing two means (Week 7).

```{r hypothesis-tests}
# Helper: run one-sample t-test and return a clean summary row
run_ttest <- function(grp, measure, pos, data_col, direction = "two.sided") {
  x   <- diff_scores %>%
    filter(group == grp, boundary_pos == pos) %>%
    pull({{ data_col }})
  tt  <- t.test(x, mu = 0, alternative = direction)
  tibble(
    Group       = grp,
    Measure     = measure,
    Position    = pos,
    N           = length(x[!is.na(x)]),
    Mean_Diff   = round(mean(x, na.rm = TRUE), 3),
    SD          = round(sd(x,   na.rm = TRUE), 3),
    t_statistic = round(tt$statistic,          3),
    df          = round(tt$parameter,          1),
    p_value     = round(tt$p.value,            4),
    CI_lower    = round(tt$conf.int[1],        3),
    CI_upper    = round(tt$conf.int[2],        3)
  )
}

# Primary tests: Post-Boundary REC and Pre-Boundary LDI for each group
results <- bind_rows(
  # H1a: Post-Boundary REC < 0 (worse recognition after boundary)
  run_ttest("Both",      "REC", "Post-Boundary", REC_diff, "less"),
  run_ttest("Item-Only", "REC", "Post-Boundary", REC_diff, "less"),
  run_ttest("Task-Only", "REC", "Post-Boundary", REC_diff, "less"),

  # H1b: Pre-Boundary LDI > 0 (better discrimination before boundary)
  run_ttest("Both",      "LDI", "Pre-Boundary",  LDI_diff, "greater"),
  run_ttest("Item-Only", "LDI", "Pre-Boundary",  LDI_diff, "greater"),
  run_ttest("Task-Only", "LDI", "Pre-Boundary",  LDI_diff, "greater")
)

ALPHA_BONF <- 0.05 / 4
results <- results %>%
  mutate(Significant = ifelse(p_value < ALPHA_BONF, "Yes *", "No"))

results %>%
  kable(caption = paste0(
    "Table 3. One-sample t-tests: difference scores vs. zero.",
    " Bonferroni-corrected α = ", round(ALPHA_BONF, 4),
    ". * = survives correction."),
    col.names = c("Group", "Measure", "Position", "N",
                  "Mean Diff", "SD", "t", "df", "p",
                  "95% CI Low", "95% CI High", "Sig.")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "bordered"),
                full_width = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#2E5FA3", color = "white") %>%
  row_spec(which(results$Significant == "Yes *"),
           background = "#E8F5E9", bold = TRUE)

write_csv(results, file.path(OUTPUT_DIR, "table3_hypothesis_tests.csv"))
```

### Figure 6: Difference Scores with Significance Annotations

```{r fig6, fig.cap="Figure 6. REC (top) and LDI (bottom) difference scores relative to Non-Boundary baseline. Horizontal dashed line at zero. Diamonds = group means. Annotated p-values from one-sample t-tests (Bonferroni corrected).", fig.height=5.5}
sig_labels <- results %>%
  mutate(
    label = paste0("p = ", p_value,
                   ifelse(Significant == "Yes *", " *", "")),
    y_pos = ifelse(Measure == "REC", 0.55, 0.55)
  )

diff_long <- diff_scores %>%
  select(participant, group, boundary_pos, REC_diff, LDI_diff) %>%
  pivot_longer(cols = c(REC_diff, LDI_diff),
               names_to  = "Measure",
               values_to = "Diff") %>%
  mutate(Measure = recode(Measure, "REC_diff" = "REC", "LDI_diff" = "LDI"))

fig6 <- diff_long %>%
  ggplot(aes(x = boundary_pos, y = Diff, colour = boundary_pos)) +
  geom_hline(yintercept = 0, linetype = "dashed", colour = "grey50") +
  geom_jitter(width = 0.18, alpha = 0.45, size = 1.3) +
  stat_summary(fun      = mean, geom = "point", shape = 18,
               size = 4, colour = "black") +
  stat_summary(fun.data = mean_se, fun.args = list(mult = 1.96),
               geom = "errorbar", width = 0.2, colour = "black") +
  geom_text(data = sig_labels,
            aes(x = Position, y = y_pos, label = label),
            colour = "black", size = 3.2, inherit.aes = FALSE) +
  scale_colour_manual(values = bp_col) +
  facet_grid(Measure ~ group) +
  labs(x     = "Boundary Position",
       y     = "Difference from Mid-Boundary Baseline",
       title = "Figure 6. Boundary Effects on REC and LDI (vs. Mid-Boundary Baseline)") +
  theme_mst +
  theme(axis.text.x = element_text(angle = 20, hjust = 1, size = 9))

fig6
ggsave(file.path(OUTPUT_DIR, "fig6_diff_scores_annotated.png"),
       fig6, width = 9, height = 5.5, dpi = 300)
```

---

# Conclusion and Future Directions

## Summary of Findings

The descriptive statistics and visualisations reveal a consistent pattern in the **Both** group:

- **Post-Boundary REC** tends to be *below* the Non-Boundary baseline, suggesting that
  items encoded immediately after an event transition are recognised less accurately.
- **Pre-Boundary LDI** tends to be *above* baseline, suggesting that items encoded
  just before a transition are represented with higher fidelity.

The one-sample t-tests begin to quantify whether these differences are statistically reliable,
though the full picture across all three groups and both boundary conditions will be clearer once
we apply additional tests in Report 2.

The encoding RT data (Figure 5) provide an important manipulation check: if participants
genuinely experienced event boundaries, responses should slow down *after* a boundary
transition — consistent with the cognitive cost of updating one's event model (Radvansky, 2017).

## Relationship to Morse et al. (2023)

The directional pattern observed here — Post-Boundary impairment for recognition, Pre-Boundary
enhancement for mnemonic discrimination — aligns with the core findings of the original paper.
Morse et al. interpreted this asymmetry in terms of hippocampal "snapshot" encoding at event
offsets: resources devoted to consolidating the ending event reduce encoding of the incoming one.

## Plans for Report 2

The following analyses are planned for subsequent reports (as the course covers the relevant methods):

- **Independent-samples t-tests** (Week 7): Compare Post-Boundary vs. Pre-Boundary directly for
  both REC and LDI within each group.
- **One-way ANOVA** (Week 9): Compare REC and LDI across the three groups to test whether
  the Both-group effects are significantly larger than Item-Only and Task-Only.
- **Lure bin analysis**: Examine whether Pre-Boundary LDI advantages are largest for high-similarity
  lures (Bin 1), where pattern separation demands are greatest.
- **Object vs. Scene breakdown**: Test whether boundary effects differ by stimulus class.

---

# References

DuBrow, S., & Davachi, L. (2013). The influence of context boundaries on memory for the sequential order of events. *Journal of Experimental Psychology: General*, *142*(4), 1277–1286.

Morse, S. J., Karagoz, A. B., & Reagh, Z. M. (2023). Event boundaries directionally influence item-level recognition memory. *Learning & Memory*.

Radvansky, G. A. (2017). Event segmentation as a working memory process. *Journal of Applied Research in Memory and Cognition*, *6*(2), 121–123.

Stark, S. M., Kirwan, C. B., & Stark, C. E. L. (2019). Mnemonic Similarity Task: A tool for assessing hippocampal integrity. *Trends in Cognitive Sciences*, *23*(11), 938–951.

Yassa, M. A., & Stark, C. E. L. (2011). Pattern separation in the hippocampus. *Trends in Neurosciences*, *34*(10), 515–525.

Zacks, J. M., Speer, N. K., Swallow, K. M., Braver, T. S., & Reynolds, J. R. (2007). Event perception: A mind-brain perspective. *Psychological Bulletin*, *133*(2), 273–293.

---

# Session Info

```{r session-info}
sessionInfo()
```
